---
title       : Full Moon and Crime Rates in the City of Chicago
subtitle    : Assignment work of Developing Data Products by JHU on Coursera
author      : T-StrawClown
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [bootstrap, quiz]            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Crimes Per Month by Moon Phases 
First lets check how crime rates look overall split by moon phase (drag to zoom, right-click to reset).
```{r permonth, message = F, error = F, warning = F, comment = NA, echo = F, results = 'asis', cache.vars=c("crimes.by_yearmonth")}
library(googleVis)
library(dplyr)
library(tidyr)
library(knitr)
library(ggplot2)

load("data/crimes.RData")

crimes.by_yearmonth <- crimes.by_day %>%
    mutate(ym = as.POSIXct(paste(year, month, "01", sep = "-"))) %>%
    #mutate(ym = as.numeric(as.POSIXct(paste(year, month, "01", sep = "-")))) %>%
    group_by(ym, moon) %>%
    summarize(cnt = sum(cnt)) %>%
    filter(!is.na(ym)) %>%
    spread(moon, cnt)

c1 <- gvisLineChart(data = crimes.by_yearmonth,
                    xvar = "ym",
                    yvar = c("New", "Waxing", "Full", "Waning"),
                    options = list(colors="['#e41a1c', '#377eb8', '#4daf4a', '#984ea3']",
                                   curveType = "function",
                                   width = 1000,
                                   height = 400,
                                   explorer = "{actions: ['dragToZoom', 'rightClickToReset']}"))
cat(c1$html$chart)
```
Doesn't seem that there is something special about full moon. BTW more detailed data can be found in my shiny application [here] (https://tstrawclown.shinyapps.io/DDPCourseProject)

--- .class #id


## Top 5 Crime Types
Here is another look. Here are top 5 types of crimes committed during full moon.
```{r columns, message = F, error = F, warning = F, comment = NA, echo = F, results = 'asis', cache.vars=c("crimes.by_typetop5", "crimes.by_type", "crimes.by_typefullmoon")}
crimes.by_type <- crimes.tidy %>%
  #filter(moon == "Full") %>%
  group_by(type, moon) %>%
  summarize(cnt = n()) %>%
  arrange(desc(cnt))
crimes.by_typefullmoon <- crimes.tidy %>%
  filter(moon == "Full") %>%
  group_by(type) %>%
  summarize(cnt = n()) %>%
  arrange(desc(cnt)) %>%
  top_n(5)
crimes.by_typetop5 <- 
  inner_join(crimes.by_type, crimes.by_typefullmoon, by = c("type" = "type")) %>%
  select(type, moon, cnt = cnt.x) %>%
  group_by(type) 
crimes.by_typetop5 <- crimes.by_typetop5 %>%
  spread(moon, cnt) %>%
  arrange(desc(Full), type)

c2 <- gvisColumnChart(data = crimes.by_typetop5,
                      xvar = "type",
                      options = list(width = 1000,
                                    height = 400,
                                    colors="['#e41a1c', '#377eb8', '#4daf4a', '#984ea3']"))
cat(c2$html$chart)
```
Doesn't look promising too...

--- &twocol w1:50% w2:50%

### Statistical Perspective
```{r, message = F, error = F, warning = F, comment = NA, echo = F, results = 'asis'}
full.ds <- subset(crimes.by_day, moon == "Full")
nonfull.ds <- subset(crimes.by_day, moon != "Full")
full.sims <- matrix(data = sample(full.ds$cnt,
                                  1000 * 1000,
                                  replace = T),
                    nrow = 1000,
                    ncol = 1000)
nonfull.sims <- matrix(data = sample(nonfull.ds$cnt,
                                     1000 * 1000,
                                     replace = T),
                       nrow = 1000,
                       ncol = 1000)
full.means <- apply(full.sims, 1, mean)
nonfull.means <- apply(nonfull.sims, 1, mean)
```

*** =left
Distribution from all data (starting from year 2009)

```{r, message = F, error = F, warning = F, comment = NA, echo = F, results = 'asis', fig.width=7, fig.height=5, fig.align='left', dpi=100}
c3 <- ggplot(data = crimes.by_day,
           aes(x = cnt, color = moon)) +
    geom_density(show.legend = F) +
    stat_density(geom = "line",
                 position = "identity") +
    #scale_color_brewer(type = "qual", guide = "legend") +
    scale_color_manual(values = c("#e41a1c", "#377eb8", "#4daf4a", "#984ea3")) +
    geom_density(data = subset(crimes.by_day, moon == "Full"),
                 aes(x = cnt),
                 color = "#4daf4a",
                 size = 1.2,
                 show.legend = F) +
    ggtitle("Density of crimes per day") +
    ylab("") + xlab("") + theme(legend.position = c(0.9, 0.8),
                                legend.background = element_blank())
print(c3)
```

*** =right
1K x 1K Bootstrap simulations (with replacement)

```{r, message = F, error = F, warning = F, comment = NA, echo = F, results = 'asis', fig.width=7, fig.height=5, fig.align='left', dpi=100}
c4 <- ggplot() +
    geom_density(data = as.data.frame(full.means),
                 aes(x = full.means),
                 color = "#4daf4a",
                 size = 1.2,
                 show.legend = F) +
    stat_density(data = as.data.frame(full.means),
                 aes(x = full.means, color = "Full"),
                 geom = "line",
                 position = "identity") +
    geom_vline(xintercept = mean(full.means) + c(-1,1) * qnorm(.975) * sd(full.means),
               linetype = "dashed",
               color = "#4daf4a") +
    geom_vline(xintercept = mean(full.means),
               color = "#4daf4a") +
    geom_density(data = as.data.frame(nonfull.means),
                 aes(x = nonfull.means),
                 color = "salmon",
                 show.legend = F) +
    stat_density(data = as.data.frame(nonfull.means),
                 aes(x = nonfull.means, color = "Other"),
                 geom = "line",
                 position = "identity") +
    geom_vline(xintercept = mean(nonfull.means) + c(-1,1) * qnorm(.975) * sd(nonfull.means),
               linetype = "dashed",
               color = "salmon") +
    geom_vline(xintercept = mean(nonfull.means),
               color = "salmon") +
    scale_colour_manual(name = "moon", 
                        values = c("Full" = "#4daf4a", "Other"="salmon"),
                        labels = c("Full", "Other"),
                        guide = "legend") +
    ggtitle("Density of simulated means of crimes per day") +
    xlab("") + ylab("") + theme(legend.position = c(0.9, 0.8),
                                legend.background = element_blank())
print(c4)
```

--- &radio

### Quiz
# The Ultimate Question

So after looking at these slides what do you think about statement that there are more crimes committed when the moon is full?

1. There is no such myth
2. _There is no statistical evidence of that_
3. It can be clearly seen that it is true
4. 42

*** .hint
No hints, sorry. You'll have to figure it out yourself

*** .explanation
Well from this presentation you can see that it is more likely that the opposite is true. Though the difference is not statistically significant.

But we both know that the true answer to the Ultimate Question is [42] (http://www.urbandictionary.com/define.php?term=42)

